### 2022-03-30 arcticgrids.tex
paper completed; ready for coauthor comments

% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
%
% You may copy this file and give it your
% article name, and enter your text.
%
%
% Step 1: Set the \documentclass
%
%

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers

%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

\draftfalse

%% Enter journal name below.
%% Choose from this list of Journals:
%
% JGR: Atmospheres
% JGR: Biogeosciences
% JGR: Earth Surface
% JGR: Oceans
% JGR: Planets
% JGR: Solid Earth
% JGR: Space Physics
% Global Biogeochemical Cycles
% Geophysical Research Letters
% Paleoceanography and Paleoclimatology
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resources Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
%
% ie, \journalname{Water Resources Research}

\journalname{Journal of Advances in Modeling Earth Systems (JAMES)}

\begin{document}

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, Marcus Lofverstrom\affil{2}, Andrew Gettelman\affil{1}}

 \affiliation{1}{National Center for Atmospheric Research, 1850 Table Mesa Drive, Boulder, Colorado, USA}
\affiliation{2}{Department of Geosciences, University of Arizona, 1040 E. 4th Street, Tucson, AZ USA}

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

%\affiliation{=number=}{=Affiliation Address=}
%(repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{Adam Herrington}{aherring@ucar.edu}

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item The transition from latitude-longitude to unstructured grids in CESM negatively impacts the surface mass balance of the Greenland Ice Sheet.
\item Two new Arctic refined mesh configurations have been developed and made available in CESM2.2
\item The Arctic meshes provide improvements to the surface mass balance over all $1^{\circ}$ grids, including latitude-longitude grids.
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT and PLAIN LANGUAGE SUMMARY
%
% A good Abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.

% The Plain Language Summary should be written for a broad audience,
% including journalists and the science-interested public, that will not have 
% a background in your field.
%
% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
% see http://sharingscience.agu.org/creating-plain-language-summary/)
%
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}

Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS). The more conventional $1^{\circ}-2^{\circ}$ uniform resolution grids systematically overestimate both the accumulation and ablation processes over the GrIS. Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids owing to their greater degrees of freedom in representing the GrIS. Two Arctic refined-meshes, containing $\frac{1}{4}^{\circ}$ and $\frac{1}{8}^{\circ}$ refinement over Greenland, are documented in this study as newly supported configurations in CESM2.2. The Arctic meshes provide substantial improvements to the simulated cloud field in the Arctic and over Greenland, resulting in a skillful representation of the accumulation and ablation processes, and a more realistic GrIS surface mass balance compared with the conventional, uniform resolution grids. As CESM is in the process of transitioning away from latitude-longitude grids, these new Arctic refined meshes serve to recover, and improve upon the representation of polar processes in CESM.

\end{abstract}

\section*{Plain Language Summary}
[ enter your Plain Language Summary here or delete this section]

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
\section{Introduction}

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high-latitudes, which are among the most sensitive regions on Earth to global and environmental change. Despite their importance, the numerical treatment of polar regions in GCMs is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}. The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a). Depending on the numerics, methods exist to suppress this instability, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain. With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past. In this study a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including brand new variable-resolution grids, are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the climate and surface mass balance of the Greenland Ice Sheet.

In the 1970's the pole problem was largely defeated through wide-spread adoption of efficient spectral transform methods in GCMs. These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g. horizontal derivatives) in the equation set can be solved for exactly. While spectral transform methods are still used in the 21st century, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems. The pole problem has thus re-emerged in contemporary climate models that use latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}. Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions. This additional damping reduces the effective resolution in polar regions such that the resolved scales are approximately the same everywhere on the grid.

An alternative approach is to use unstructured grids, which allow for more flexible grid structures that permit quasi-uniform grid spacing globally and eliminates the pole-problem entirely (e.g., Figure~\ref{fig:uni-grids}c). This grid flexibility also permits variable-resolution or regional grid refinement (e.g., Figure~\ref{fig:vr-grids}). Grids can be developed with refinement over polar regions that could in principle make up for any loss in polar resolution in transitioning away from latitude-longitude grids (e.g., Figure~\ref{fig:vr-grids}), although this comes at the cost of a smaller CFL-limiting time-step in the refined region (the CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space). Unstructured grids also scale more efficiently on parallel systems than latitude-longitude grids, likely resulting in a greater prevalence of unstructured grids as computing power continued to increase over time.

The meteorology and climate of the Arctic is characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}. For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees \cite{JW2006QJR,IPCCAR5}, mesoscale Polar Lows are not well resolved at these resolutions. These mesoscale systems are prevalent during the cold season and produce gale-force winds that can induce large heat and moisture fluxes through the underlying sea-ice/ocean interface. The Arctic also contains the Greenland Ice Sheet (hereafter denoted as \textit{GrIS}). While it blankets the largest island in the world (Greenland), many of the processes that control the GrIS annual surface mass balance (the integrated sum of precipitation and runoff) are only partially resolved at typical GCM resolutions. For example, GrIS precipitation is concentrated at the ice-sheet margins, where orographic precipitation is generated by steep topographic slopes. GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically 10s to 100\,km wide and confined to low-level areas or regions with limited precipitation. GCMs struggle to resolve the magnitude and extent of these features \cite{P2010CC,VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.

The goal of this study is to characterize the representation of high-latitude regions using the spectral-element and finite-volume dycores in CESM2.2, as these models treat the high-latitudes, e.g., the pole-problem, in different ways. The manuscript is laid out as follows: Section~\ref{sec:methods} consists of documentation of the grids, dycores and physical parameterizations used in this study. The Arctic refined grids were developed by the authors, and this section serves as their official documentation in CESM2.2. Section~\ref{sec:methods} also contains a description of the experiments along with the datasets and methods for evaluating the model simulations. Section~\ref{sec:results} contains the results of the experiments, followed by Section~\ref{sec:conclusions} that provides a general discussion and conclusions.

\begin{figure}[t]
\begin{center}
\begin{tabular}{cc}
         \includegraphics[width=60mm]{figs/grid-f19.pdf}&
         \includegraphics[width=60mm]{figs/grid-f09.pdf}\\
         \includegraphics[width=60mm]{figs/grid-ne30pg2.pdf}&
         \includegraphics[width=60mm]{figs/grid-ne30pg3.pdf} \\
\end{tabular}
\end{center}
\caption{Computational grids for the uniform $1^{\circ}-2^{\circ}$ grids in this study.}
\label{fig:uni-grids}
\end{figure}

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=60mm]{figs/grid-ARCTIC.pdf}&
         \includegraphics[width=60mm]{figs/grid-ARCTICGRIS.pdf} \\
\end{tabular}
\end{center}
\caption{Spectral-element grid for the variable-resolution ARCTIC grid in this study. Note that this is not the computational grid; each element has $3\times3$ independent grid points.}
\label{fig:vr-grids}
\end{figure}

\section{Methods}\label{sec:methods}
\subsection{Dynamical cores}

The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 \cite<CAM;>{CAM63}, supports a number of different atmospheric dynamical cores. These include dycores using latitude-longitude grids, such as finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models. The EUL dycore is the oldest dycore in CAM, and the least supported of all the dycores. FV3 is the newest dycore in CAM, but it was not fully incorporated at the time this work commenced; both the EUL and FV3 dycores are omitted from this study. As such, the results presented in this study are comparing the performance of the SE and FV dycores.

\subsubsection{Finite-volume (FV) dynamical core}

The FV dycore is a hydrostatic model that integrates the equations of motion using a finite-volume discretization on a spherical latitude-longitude grid \cite{LR1997QJR}. The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to Eulerian reference grid in the vertical \cite{L2004MWR}, using a hybrid-pressure vertical coordinate. Hyperviscous damping is applied to the divergent modes while Laplacian damping is applied to momentum in the top few layers, referred to as a \textit{sponge layer} \cite{L2011IJHPC}. A polar filter is used to avoid computational instability due to the convergence of the meridians, allowing for a more practical time-step. It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the poleward direction \cite{ST1995GEOS}.

\subsubsection{Spectral-element (SE) dynamical core}

The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}. The computational domain is a cubed-sphere grid tiled with quadrilateral elements (e.g., Figure~\ref{fig:vr-grids}). Each element contains a fourth order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre (GLL) quadrature points. This results in 16 GLL nodal points within each element, with 12 of the points lying on the (shared) element boundary. Communication between elements happens via the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries that reconciles overlapping nodal values and produces a continuous global basis set.

As with the FV dycore, the dynamics evolve in floating Lagrangian layers that are subsequently mapped to an Eulerian reference grid. A dry mass vertical coordinate was more recently implemented for thermodynamic consistency with condensates \cite{LetAl2018JAMES}. The 2D dynamics have no implicit dissipation and so hyperviscosity operators are applied to all prognostic variables to remove spurious numerical errors \cite{DetAl2012IJHPCA}. Laplacian damping is applied in the sponge layer.

The SE dycore supports regional grid refinement via its variable-resolution configuration, requiring two enhancements over uniform resolution grids. (1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing. A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements. (2) The topography boundary conditions need to be smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size.

For spectral-element grids with quasi-uniform grid spacing, a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) is used instead \cite<>{LTOUNGK2017MWR}. CSLAM has improved tracer property preservation and accelerated multi-tracer transport. It uses a separate grid from the spectral-element dynamics, through dividing each element into $3\times3$ control volumes with quasi-equal area. The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the default SE dycore in which the physics are evaluated at the GLL nodal points \cite{HL2018MWR}.

\subsection{Grids}

Six grid are evaluated in this study (Table~\ref{tbl:table1}). The FV dycore is run with $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b). The $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid is also run, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics. An additional $1^{\circ}$ CAM-SE-CSLAM grid is run, but with the physical parameterizations computed on a grid that contains $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.

Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic (Figure~\ref{fig:vr-grids}). This manuscript serves as the official documentation of these grids. The Arctic meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}). The $ARCTIC$ grid is a $1^{\circ}$ grid with $\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region. The $ARCTICGRIS$ grid is identical to the $ARCTIC$ grid, but contains an additional patch covering the big island of Greenland with $\frac{1}{8}^{\circ}$ resolution.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lccccc}
   \hline
   grid name & dycore & $\Delta x_{eq}$ (km) & $\Delta x_{refine}$ (km) & $\Delta t_{phys}$ (s) \\ 
   \hline
   $f19$ & FV & 278 & - &1800 \\
   $f09$ & FV & 139 & - &1800 \\
   $ne30pg2$ & SE-CSLAM & 167 & - & 1800 \\
   $ne30pg3$ & SE-CSLAM & 111 & - & 1800 \\
   $ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 \\
   $ARCTIC$ & SE & 111 & 28 & 450 \\
   $ARCTCIGRIS$ & SE & 111 & 14 & 225 \\
 \hline
 \end{tabular}
  \caption{Grids and dycores used in this study. $\Delta x_{eq}$ refers to average equatorial grid spacing, $\Delta x_{refine}$ refers to grid spacing in the refined region (if applicable) and $\Delta t_{phys}$ refers to the physics time-step. The dycore abbreviation FV refers to the finite-volume dycore, SE the spectral-element dycore and SE-CSLAM the spectral-element dycore w/ CSLAM tracer advection.}
 \label{tbl:table1}
 \end{table*}

The physics time-step should depend on grid resolution. Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step is reduced to avoid large time truncation errors \cite{HR2018JAMES}. The $ARCTIC$ and $ARCTICGRIS$ grids are run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in the standard $1^{\circ}-2^{\circ}$ grids (Table~\ref{tbl:table1}).

All grids and dycores in this study use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$. However, any grid or dycore can in principle be run with a higher-top and/or with finer vertical resolution.

\subsection{Physical parameterizations}

The CAM6.3 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{CAM63} is used in all simulations in this study. CAM version 6 physics and later is most notably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme. CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory in representing cloud ice for improved cloud-aerosol interactions. Deep convection is parameterized using a convective quasi-equilibrium, bulk plume mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}. PBL form drag is modeled after \cite{BBW2004QJRMS} and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale.

Initial simulations with the $ne30pg3$ spectral-element grid produced weaker shortwave cloud forcing relative to the tuned up finite-volume dycore. All runs with the spectral-element dycore have two CLUBB parameter changes in order to provide a more realistic cloud forcing and top-of-atmosphere radiation balance. These are CLUBB's $gamma$ parameter, reduced from 0.308 to 0.270, and c14, reduced from 2.2 to 1.6. Briefly, the $gamma$ parameter scales the width of the sub-grid distribution of vertical velocity, and c14 controls the strength of the damping term in the equation for the horizontal component of turbulent kinetic energy. For a thorough explanation of how CLUBB parameters impact the simulated climate, the reader is referred to \cite{GETAL2015JAMES}.

\subsection{Experimental design}

All grids and dycores are run using an identical transient 1979-1998 AMIP-style configuration, with prescribed monthly SST/sea-ice after \cite{CESMSST}. This configuration refers to the $FHIST \ compset$ and runs out of the box in CESM2.2.

The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study. The SMB is the sum of the mass accumulation term,  i.e., precipitation, and the mass sink term, ablation. Ablation can be expressed as evaporation/sublimation plus total runoff, with runoff being a combination of liquid precipitation and snow and ice melt. Not all liquid precipitation runs-off the ice sheet; rain can penetrate pore spaces in the snowpack/firn layer and freeze, representing an accumulation of mass. These processes are represented by different components in CESM, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.

CLM runs on the same grid as the atmosphere, but also uses a downscaling technique to account for sub-grid variability in SMB. In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (EC), weighted by their respective area fractions at each EC, which is derived from a high resolution GrIS elevation dataset. The near surface air temperature, humidity and air density are calculated at each EC using an assumed lapse rate and the elevation difference from the grid mean, and the precipitation rates from CAM are repartitioned into solid or liquid based on the temperature of the EC. Ice accumulation is modeled as a capping flux, or snow in excess of a 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of ice. A unique surface energy balance and SMB is computed for each EC. Integrating over all ECs using the area weights provides a more accurate SMB. For a more detailed description of how the SMB is computed in CESM, the reader is referred to \citeA{CISM1,SETAL2019TC,KETAL2020JAMES}.

Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up by forcing CLM in standalone mode, cycling over a 20 year $ARCTIC$ $FHIST$ run for about 500 years. The uniform resolution grids are all initialized with an SMB from an existing $f09$ spun-up initial condition.

\subsection{Observational Datasets}

Several observational datasets are used in this study to gauge the performance of the simulations. A list of the datasets used in this study are provided in Table~\ref{tbl:table2}. Several of these products are near-global gridded datasets commonly used to evaluate GCMs. Surface mass balance datasets were initially gathered from several sources. RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA40, ERA interim and ERA5 renalysis products at its domain boundaries. The RACMO simulations have been shown to perform skillfully against observations and is often used as an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}.

In-situ SMB (snow pit and ice cores) and radar accumulation datasets (e.g., IceBridge) are conveniently maintained in The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit}. However, these point-wise measurements are difficult to compare to model output spanning several different grids, and would be more useful if the surface mass balance from the individual EC's were available from the model output, which they are not. A nearest neighbor technique was used to do an initial analysis, and indicated that the model biases are similar to those computed using the RACMO datasets. Owing to the uncertainty of comparing gridded fields to point-wise measurements, and the lack of additional information it adds with regard to model biases, these datasets were omitted from the analysis in this study.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lcccc}
   \hline
   data product & years used in this study & resolution & citation \\ 
   \hline
   ERA5 & 1979-1998 & $\frac{1}{4}^{\circ}$ & \citeA{ERA5} \\
   CERES-EBAF ED4.1 & 2003-2020 & $1^{\circ}$ & \citeA{CERES-EBAF} \\
   CALIPSO-GOCCP & 2006-2017 & $1^{\circ}$ & \citeA{CALIPSO-GOCCP} \\
   RACMO2.3 & 1979-1998 & 11 km & \citeA{NETAL2015TC} \\
   RACMO2.3p2 & 1979-1998 & 5.5 km & \citeA{NETAL2019SCIENCE} \\
 \hline
 \end{tabular}
  \caption{Description of observational datasets used in this study.}
 \label{tbl:table2}
 \end{table*}

\subsection{SMB Analysis}\label{sec:SMB}

We seek to integrate components of the SMB over a GrIS ice mask and diagnose their contributions to the mass budget of the GrIS. However, the ice masks vary across all the grids, especially in comparison to the RACMO3.2 ice mask which is about $3\%$ smaller than the reference dataset (Figure~\ref{fig:grisdx}). The area errors in Figure~\ref{fig:grisdx} may not seem large, but even order $1-2\%$ percent ice mask area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}. CLM's dataset creation tools generates the model ice mask by mapping a raw high resolution dataset containing the GrIS boundary conditions, to the target grid using the Earth System Modeling Framework \cite<ESMF;>{ESMF} first-order conservative remapping algorithm. The figure suggests the mapping errors are less than $1.5\%$ across just the model grids.

In this study we have taken a common ice mask approach by mapping all model fields to the lowest resolution grids and integrating over their respective low resolution ice masks. A low resolution common ice mask is a conservative decision and is largely because we seek to use first-order remapping algorithms for mapping fields to the common ice mask, which is not reliable in the downscaling direction, i.e., mapping to a higher resolution grid than the source grid. Two remapping algorithms are used; ESMF first-order conservative and TempestRemap \cite{TempestRemap} high-order monotone algorithm. Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$ grids, the lowest resolution grid for each grid coordinate type. In all, each integrated quantity is evaluated (at most) four times per grid to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.

We consider two different approaches for integrating mapped fields over the common ice mask.
\begin{enumerate}
\item Map the patch level quantities, or the state over the ice fractional component of the grid cell, to the common grid, and integrate the mapped field over the common ice mask. \label{lbl:label1}
\item Map the grid cell mean quantities to the common grid, and integrate the mapped field over the common ice sheet. \label{lbl:label2}
\end{enumerate}

Method~\ref{lbl:label1} introduces errors that is related to differences in the ice mask between the source and target grid. Since the lower resolution ice masks tend to have larger GrIS areas (Figure~\ref{fig:grisdx}), we expect these errors to inflate the resulting integrals. Method~\ref{lbl:label2} assumes the patch level values are identical to the grid mean values. This has the effect of homogenizing ice mask specific states over the entire grid cell. This is especially detrimental to the ice melt quantity, since ice melt can only occur over the ice mask. The accumulation term (combined solid plus liquid precipitation), however, should not be adversely impacted by method~\ref{lbl:label2} since it is not directly tied to the ice mask, but instead controlled by larger resolved scales. One could also argue that method~\ref{lbl:label1} is less accurate for the accumulation term, since it does not map to the new ice mask as cleanly as in method~\ref{lbl:label2}.

Method~\ref{lbl:label2} may be necessary in situations where only the grid mean values are available, and the patch level values cannot be reconstructed. This is unfortunately the case for the RACMO datasets; ice melt cannot be reconstructed at the ice patch level. Method~\ref{lbl:label2} is therefore used in this study, and we should expect a dissipative effect when diagnosing the melt rates. As we will show later, this method still preserves the relative differences between the different grids and in comparison with the RACMO datasets.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_grisres.pdf}
\end{center}
\caption{The spatial properties of the GrIS as represented by different grids in this study. (Left) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
\label{fig:grisdx}
\end{figure}

\section{Results}\label{sec:results}

\subsection{Tropospheric temperatures}

Before delving into the simulated characteristics of the Arctic, the global mean differences between the various grids and dycores are assessed. Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between the uniform resolution grids and dycores. The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere. This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \citeA{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB. The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the low and mid-latitudes in the $f09$ grid.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-lores.pdf}
\end{center}
\caption{1979-1998 annual mean (left column) temperature and (right column) CLUBB temperature tendencies in zonal mean height space, expressed as differences between the various $1^{\circ}-1^{\circ}$ grids. The thick grey and magenta lines are the tropopause for control run and the test run, respectively.}
\label{fig:dT-lores}
\end{figure}

As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so a modest, resolution-like sensitivity occurs in which $ne30pg3$ is warmer than $f09$ (Figure~\ref{fig:dT-lores}). The stratosphere has a uniquely different response, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes. Figure~\ref{fig:dT-lores} also shows differences in temperature between $ne30pg3$ and $ne30pg2$, which are small, although there is a slight warming near the tropopause at high latitudes. This is consistent with the similar climates found between these grids in \citeA{HETAL2019JAMES}.

Comparing the variable-resolution grids to the uniform resolution grids is complicated because we simultaneously increase the grid resolution and reduce the physics time-step, both which noticeably impact the solution \cite{W2008TELLUS}. An additional $ne30pg3$ simulations is run with the physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$. Figure~\ref{fig:dT-hires} shows the change in the climatological summer temperatures in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$. A similar warming response to increasing resolution occurs when the time-step is reduced, and the mechanism is similar in that the shorter time-step facilitates greater condensational heating by CLUBB. Figure~\ref{fig:dT-hires} shows the difference in climatological summer temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid. The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the comparison.

Figure~\ref{fig:dT-hires} indicates that the $ARCTICGRIS$ grid is a lot warmer than the $ARCTIC$ grid in the Arctic. Part of this may be due to the shorter physics time-step; a similar physics-time-step sensitivity in the $ne30pg3^{*}$ comparison is evident in the CLUBB tendencies in the $ARCTCGRIS$ grid. However, the temperature response is disproportionately large relative to the change in CLUBB tendencies. This summer warming appears to be a result of variations in the stationary wave pattern, with anomalous southerly winds occurring to the west of Greenland (not shown). This dynamic response is extremely interesting because other than the physics time-step, the only difference between $ARCTICGRIS$ and the $ARCTIC$ grid is the doubling of resolution of Greenland. This behavior will be explored more fully in a future study.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-hires.pdf}
\end{center}
\caption{As in Figure~\ref{fig:dT-lores} but for the short-time step experiment and the variable-resolution grids. The fields plotted are for the climatological northern hemisphere summer. The focus on summer is because that is when the resolution response is largest, and the refined regions are located in the northern hemisphere.}
\label{fig:dT-hires}
\end{figure}

It's useful to understand summer temperature biases due to its control on ice/snow melt over the Greenland Ice Sheet \cite<GrIS;>{O2001JAM}. Figure~\ref{fig:dThyps} shows the 1979-1998 lower troposphere summer temperature bias relative to ERA5. It is computed from the 500 hPa-1000 hPa geopotential thickness, solving for the layer mean virtual temperature using the hypsometric equation. The results generally track with the analysis of the zonal mean height plots; increasing resolution from $f19$ to $f09$ leads to a warmer climate, and the $1^{\circ}$ spectral-elements grids are warmer than the finite-volume grids. The summer temperatures in the finite-volume grids are persistently colder than ERA5, whereas the $1^{\circ}$ spectral-element grids are not as cold, and are actually warmer than ERA5 at very high-latitudes, north of $75^{\circ}$. All grids illustrate a north-south gradient in bias over Greenland, in which the summer temperature bias becomes more positive on the northern side of the ice sheet. This pattern is also evident in the 2m summer temperature bias over Greenland (not shown).

The $ARCTIC$ grid has similar summer temperatures to the $1^{\circ}$ spectral-element grids, but it is a bit warmer over northern Eurasia and the North Pole. An anomalous cooling patch forms to the west of Greenland, centered over Baffin Island. The $ARCTICGRIS$ grid is warmer than the $ARCTIC$ grid over most of the Arctic, but maintains a similar spatail pattern of summer temperature bias as in the $ARCTIC$ grid.

Some of these temperature anomalies may be related to summer shortwave cloud forcing differences across the different grids and dycores. Figure~\ref{fig:SWCF} shows the summer short-wave cloud forcing bias in the runs, using the CERES-EBAF product. All the uniform $1^{\circ}-2^{\circ}$ grids have similar biases, with the clouds reflecting 20-40 W/m2 too much shortwave radiation over a wide swath of the Arctic, primarily over the land masses. There's also a halo of low cloud forcing bias around the oceanic perimeter of Greenland. The $ARCTIC$ grid has much smaller cloud forcing biases over the Arctic land masses, although still too reflective, whereas the $ARCTCGRIS$ grid vastly improves the cloud forcing bias over Eurasia, and improves the bias over N.America compared to the $ARCTIC$ grid. In both variable-resolution grids, the halo of too weak cloud forcing bias around the perimeter of Greenland is absent.

While the summer cloud forcing biases are consistent with the summer temperature biases in Figure~\ref{fig:dThyps} --regions where clouds are too reflective coincide with regions that are too cold-- it is not clear whether the cold biases are caused by the cloud biases, or whether the cold biases amplify the cloud forcing bias.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_contours_diffERA5_Thyps.pdf}
\end{center}
\caption{1979-1998 lower troposphere, northern hemisphere summer virtual temperature biases, computed as the difference from ERA5. Lower troposphere layer mean virtual temperature is derived from the 1000 hPa - 500h Pa geopotential thickness, using the hypsometric equation. Differences are computed after mapping the ERA5 data to the finite-volume grids since the geopotential field is only available on the output tapes in the spectral-element runs that have been interpolated to the $f09$ grid, inline.}
\label{fig:dThyps}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_contours_diffCERES_SWCF.pdf}
\end{center}
\caption{1979-1998 northern hemisphere summer shortwave cloud forcing bias, relative to the CERES-EBAF gridded dataset. Differences are computed after mapping all model output to the $1^{\circ}$ CERES-EBAF grid.}
\label{fig:SWCF}
\end{figure}

\subsection{Shortwave radiation over Greenland}

In addition to summer temperatures, shortwave radiation is also an important determinant of snow/ice melt. Figure~\ref{fig:FSDS} shows the summer incident shortwave radiation bias at the surface, zoomed in over Greenland. The top panel computes the bias using the CERES-EBAF dataset, and the bottom panel using RACMO2.3p2 dataset.
This halo of excessive incident shortwave radiation around the coasts of Greenland is apparent in both datasets, consistent with the shortwave cloud forcing biases in Figure~\ref{fig:SWCF}.

The interior of the ice sheet receives too little shortwave radiation in the coarser grids. In the variable-resolution grids, both the interior deficit in shortwave and the excessive shortwave around the oceanic perimeter of Greenland are improved. This suggests that the coarse grids clouds are too thick in the interior of Greenland, and too thin around the perimeter of Greenland, and that increasing horizontal resolution balances out these biases. This is consistent with total summer cloud fraction bias, computed from the CALIPSO-GOCCP cloud dataset (Figure~\ref{fig:prect}). Note that total cloud fraction characterizes the cloud field at all vertical levels, but attenuates any changes arising from any single layer due to the maximum overlap assumption used to compute this quantity. Despite the attenuated signal, the total cloud fraction does indicate a reduction in cloud coverage in the interior, and an increase in cloudiness about the oceanic perimeter, in the variable-resolution grids.

The agreement of the cloud biases in and around Greenland from multiple independent datasets indicates this is a robust feature of the coarser grids. The reduction of these cloud biases in the variable-resolution grids suggests they are a result of insufficient horizontal resolution in the coarse grids.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_contours_diffCERESdiffRACMO_FSDS.pdf}
\end{center}
\caption{1979-1998 northern hemisphere summer, incident shortwave radiation bias, computed as the difference (top) from CERES-EBAF, and (bottom) RACMO2.3p2 dataset. The differences in the top panel are found by mapping the model output to the $1^{\circ}$ CERES-EBAF grid, and differences on the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids. Note that the averaging period for the CERES-EBAF panels, 2003-2020, is different from the averaging period for the model results.}
\label{fig:FSDS}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_contours_diffCERESdiffRACMO_CLOUD_PRECIP.pdf}
\end{center}
\caption{1979-1998 northern hemisphere summer (top) total cloud fraction bias, relative to the CALIPSO-GOCCP dataset, and (bottom) precipitation rate bias, relative to the RACMO2.3p2 dataset. The CALIPSO-GOCCP differences are computed after mapping all model output to the $1^{\circ}$ grid, whereas the RACMO differences are computed after mapping the RACMO dataset to the individual model grids. Note that the averaging period for the CALIPSO-GOCCP panels, 2006-2017, is different than the model averaging period. \color{red}{ARH - still trying to fix the layout of this figure so the label bars aren't on top of the panels}.}
\label{fig:prect}
\end{figure}

\subsection{Greenland surface mass balance}

The accuracy of the simulated SMB is expected to be sensitive to grid resolution. Figure~\ref{fig:prect} shows the average grid spacing over the Greenland Ice Sheet (GrIS) in all six grids in this study. The $ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $ARCTICGRIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO2.3 grid. The $ne30pg3$ grid has an average $\Delta x=111.2~km$, which is substantially more coarse than the $f09$ grid, with an average $\Delta x=60~km$. This is interesting because $ne30pg3$ and $f09$ have similar average grid spacing over the entire globe, and comparable computational costs, but due to the convergence of meridians the finite-volume model has enhanced resolution over GrIS. The $ARCTIC$ grid has an average grid spacing of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models (whereas the $ARCTICGRIS$ grid is about twice as expensive as the $ARCTIC$ grid).

The summer climatological mean precipitation bias over GrIS is shown in Figure~\ref{fig:prect}, expressed as the fractional difference from the RACMO2.3p2 solution. It is evident that the coarse $1^{\circ}-2^{\circ}$ grids have large, positive biases centered over the southern dome. The $ARCTIC$ run improves this bias substantially, and the $ARCTICGRIS$ run improves this bias further. This suggests the southern dome bias is due to inadequate horizontal resolution, which is consistent with the original GrIS variable-resolution experiments in \cite{VETAL2018TC}.

Large accumulation rates in GrIS result from synoptic systems arriving from the south. These systems are orographiclly lifted at the ice sheet margin, especially the steep slopes in southeast Greenland, concentrating large amounts of precipitation about the ice margin. At lower horizontal resolutions the topography is too smooth and anomalous moisture penetrates inland, erroneously dumping precipitation onto the southern dome. The ability of the variable-resolution grids to more accurately simulate the orographic precipitation process in Greenland is consistent with all the cloud results up to this point. As the precipitation centers move from the interior towards the margins, and even out over the ocean with increasing resolution, the cloud decks should, and are, moved accordingly. Figures~\ref{fig:SWCF}, \ref{fig:FSDS}, \ref{fig:prect} clearly illustrate a reduction in cloud amount in the interior and an increased cloudiness around the oceanic perimeter of Greenland, in comparing the uniform $1^{\circ}-2^{\circ}$ runs to the variable-resolution runs.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lccccc}
   \hline
   grid name & accumulation & total melt & runoff & sublimation & SMB \\ 
   \hline
   $RACMO$ & 768.5 (733.5) & -347.2 (-436.4) & -221.7 (-258.5) & -36.5 (-38.8) & 510.3 (436.2) \\
   \hline
   $f19$ & 882.5 (913.5) & -440.3 (-546.5) & -283.7 (-284.3) & -36.6 (-37.5) & 562.2 (591.7) \\
   $f09$ & 874.8 (882.1) & -418.4 (-482.3) & -255.0 (-212.3) & -38.1 (-37.4) & 581.7 (632.4) \\
   $ne30pg2$ & 1000. (973.4) & -549.4 (-647.3) & -383.9 (-347.0) & -33.4 (-32.1) & 582.7 (594.3) \\
   $ne30pg3$ & 934.9 (909.3) & -568.8 (-686.7) & -356.2 (-330.1) & -34.4 (-32.6) & 544.3 (546.6) \\
   $ARCTIC$ & 795.9 (818.6) & -367.3 (-436.8) & -208.9 (-194.2) & -44.1 (-43.9) & 542.9 (580.5) \\
   $ARCTICGRIS$ & 708.7 (747.3) & -471.6 (-610.4) & -261.1 (-307.8) & -50.7 (-51.8) & 396.9 (387.7) \\
 \hline
 \end{tabular}
  \caption{1979-1998 surface mass balance of the Greenland Ice Sheet in Gt/yr. Values shown are using the common ice mask approach described in the methods section, whereas values in parentheses are from integrating over the native grid and ice mask.}
 \label{tbl:table3}
 \end{table*}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_tseries_GRIS.pdf}
\end{center}
\caption{Time-series of annual (solid+liquid) precipitation (top) and annual runoff (bottom) integrated over the Greenland Ice Sheet for all six simulations and compared to the RACMO datasets. The time-series were generated using the common ice mask approach, which results in up to 4 ensembles, with the mean value given by the solid line and shading spanning the extent of the ensemble members.}
\label{fig:tseries}
\end{figure}

Table~\ref{tbl:table3} shows the 1979-1998 climatological SMB components in all the grids, and compared with RACMO. Note the values in the table averages over the ensemble of common ice masks and regridding methods described in section~\ref{sec:SMB}. The RACMO values are averages over both RACMO datasets (Table~\ref{tbl:table2}) using the same common ice mask approach. The SMB is computed from the individual components as:
\begin{equation}
SMB = accumulation + runoff + sublimation \label{eq:SMB}
\end{equation}
and all terms have consistent sign conventions (positive values contribute mass, negative values reduce mass). Note that this glaciological SMB differs from the internally computed SMB in CLM; see \citeA{KETAL2020JAMES} for an explanation of the differences.

Table~\ref{tbl:table3} also contain's (in parentheses) the SMB components derived from evaluating the integrals on each model's native grid and ice mask. Of note is the large reduction in ice/melt rates using the common ice mask approach compared with that computed on the native grid. This illustrates the dissipative impact of the common ice mask approach discussed in section~\ref{sec:SMB}. The errors are large because it is mixed ice/ice-free grid cells straddling the ice sheet margins that are more prone to errors, which are commonly in the ablation zone and have large melt rates. Contrast this with the differences in integrated precipitation, where the differences between the native approach and common ice mask approach is much less. As discussed in the methods, this is because the combined solid/liquid precipitation rates are not directly tied to the ice mask, whereas ice melt is exclusively tied to the ice mask.

Figure~\ref{fig:tseries} shows time-series of annually integrated precipitation and snow/ice melt over the GrIS in the various different grids and dycores, with both versions of RACMO shown in black. The 1979-1998 climatological mean values, tabulated in Table~\ref{tbl:table3}, are also provided as circles on the right side of the panels. The uniform $1^{\circ}-2^{\circ}$ grids all show a distinctive positive bias in precipitation due to this over-prediction of interior precipitation rates. The variable-resolution grids have the smallest precipitation biases, providing a comparable solution to RACMO. The $f19$ and $f09$ perform similarly, with +110 Gt/yr bias, whereas $ne30pg3$ is biased by about +165 Gt/yr and $ne30pg2$, +230 Gt/yr. The results suggest that uniform resolution spectral-element grids have larger biases than the finite-volume grids, consistent with spectral-element grids having a coarser representation of GrIS (Figure~\ref{fig:grisdx}).

The combined annual snow/ice melt shown in the bottom panel of Figure~\ref{fig:tseries} indicates that the $ARCTIC$ grid simulates the most realistic melt rates, with all other grids tending to have larger melt rates than RACMO. The $ARCTICGRIS$ grid over predicts melting by about 125 Gt/yr. This is likely due to an anomalously warm lower troposphere during the summer, relative to the $ARCTIC$ run (Figure~\ref{fig:dThyps}). The $f19$ and $f09$ melting rates are improved over $ARCTICGRIS$, overestimating melt by only 70-90 Gt/yr. The spectral-element grids have the largest positive melt bias, between 200-220 Gt/yr. It is more difficult to attribute these differences to resolution alone, since the finite-volume grids have colder summer temperatures than the uniform resolution spectral-element grids.

To understand the regional behavior of the SMB components, Figure~\ref{fig:basin} shows the precipitation and combined snow/ice melt integrated over the basins defined by \citeA{RM2012GRL}. The uncertainty due to differences in basin area is larger than for GrIS wide integrals, owing to the differences in basin boundaries as represented by the common ice masks, which can be seen in the $f19$ and $ne30pg2$ panels of Figure~\ref{fig:melt}. Nonetheless, the regional totals in Figure~\ref{fig:basin} correctly show the southeast and southwest basins have the largest accumulation totals. All basins exhibit a monotonic reduction of accumulation rates with increasing grid resolution, with some exceptions. The $ARCTCIGRIS$ grid simulates less precipitation than RACMO in the central-east and southeast basins, and is closest of all grids to the RACMO precipitation amounts in the large southwest basin.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_tseries_BASIN.pdf}
\end{center}
\caption{1979-1998 basin integrated components of the SMB; (top) precipitation, (middle) ice/snow melt and (bottom) ice/snow melt estimated from the PDD method. Whiskers span the max/min of the 4 ensemble members generated from the common ice mask approach. Basin definitions are after \citeA{RM2012GRL}, and are found on the common ice masks using a nearest neighbor approach, and shown in Figure~\ref{fig:melt}.}
\label{fig:basin}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_contours_diffRACMO_melt.pdf}
\end{center}
\caption{1979-1998 ice/snow melt biases (in mm/yr) relative to RACMO2.3p2, evaluated on the native model grids. The \citeA{RM2012GRL} basin boundaries are shown in grey for each model grid. Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $f19$ and $ne30pg2$ panels, in computing the basin-scale integrals. Blue lines in the $f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}.}
\label{fig:melt}
\end{figure}

The basin integrated melt rates in Figure~\ref{fig:basin} show a dycore dependent-type pattern. The uniform-resolution spectral-element grids have the largest positive bias in melt rates, in all basins. The $ARCTICGRIS$ grid is a close second, while the finite-volume grids have systematically smaller melt-rates. The ``second-place" standing of $ARCTICGRIS$ is somewhat unexpected as this grid has the warmest lower troposphere summer temperatures (Figure~\ref{fig:dThyps}) and incident shortwave radiation (Figure~\ref{fig:FSDS}), yet it has less melting than the uniform-resolution spectral element grids.

Lower troposphere temperature is not a strict proxy for melting; e.g., it may not capture microclimate effects as a result of a better representation of the low-elevation ablation zones. A more accurate proxy for the impact of temperature on melting is positive degree-days \cite<PDD;>{B1984Z}. PDD is a metric that accumulates the near surface temperature in $^{\circ}C$ for any day in which the temperature is above freezing. PDD is non-linear in mean monthly temperature \cite{R1991P}, and we compute it using monthly mean 2-meter temperature using the method of \citeA{CG2005JG}, assuming a fixed monthly mean standard deviation of 3$^{\circ}C$, and a degree-day factor of 5 $mm \ d^{-1}$ $^{\circ}C^{-1}$.

The basin integrated PDD melt estimate is shown in Figure~\ref{fig:basin}. One interesting feature is that in the big southeast and southwest basins (and all the other western basins), the $ne30pg3$ grid has larger PDD based melt than the $ARCTCGRIS$ grid. The southwest basin also predicts a rather large PDD based melt in the finite-volume grids, relative to the $ARCTCIGRIS$ grid. The PDD plots indicate that the near surface temperatures expected to contribute to melt is not well approximated by the summer lower troposphere temperatures in Figure~\ref{fig:dThyps}.

Figure~\ref{fig:melt} shows the biases in the combined ice/snow melt as map plots. This plot is quite revealing as it indicates that the largest melt biases occur on the southeast and northwest coast, where large coarse grid cells overlap with the ocean. One possibility is that these problematic grid cells are situated at lower elevations then the ice sheet surface exists in nature, leading to too warm of a state and overestimating the melt. Figure~\ref{fig:ztrans} shows the representation of the surface of the ice sheet along two transects in the different grids, compared to the high resolution dataset used to generate CAM topography boundary conditions \cite{GMTED2010,gmdd-8-4623-2015}. The two transects are shown in Figure~\ref{fig:melt}; the infamous "k-transect" in southwest Greenland and a transect extending from the central dome down to the highly productive Kangerlussuaq glacier on the southeast coast. The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect. The $f09$ grid is a bit of an exception; the grid cells become very narrow in the meridional direction at high latitudes and a larger number of grid cells populate the east-west transect. The variable-resolution grids are more skillful at reproducing the steep margins of the ice sheet, capturing the parabolic shape of the GrIS margins.

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=130mm]{figs/temp_zsects.pdf}
\end{tabular}
\end{center}
\caption{Model surface elevation along the (top) k-transect, and (bottom) a transect spanning the central dome, down to the Kangerlussuaq glacier in southeast Greenland, for all model grids. The reference surface (GMTED) is a 1 km surface elevation dataset used for generating the CAM topography boundary conditions. }
\label{fig:ztrans}
\end{figure}

While the transects in Figure~\ref{fig:ztrans} only provide two anecdotes, they indicate that the ice sheet surface is not systematically lower than the true surface in the coarse grids, in the ablation zones. Rather the smoothing of the raw topography, necessary to prevent the model from instigating grid-scale modes, flattens the ice sheet, causing the lower-elevation ablation zones to extend further outside the true ice sheet margin, and are situated higher than the actual ice surface. The $f19$ grid has the smoothest topography since it's dynamics are coarsest (whereas $f09$, $ne30pg2$ and $ne30pg3$ use identical smoothing), and has the flattest ice sheet. This suggests that if anything, coarser models will tend to elevate the ablation zones and depress melt rates.

Figure~\ref{fig:ztrans} also labels the ice margin boundary, which nicely illustrates that the ablation zone occurs over very narrow horizontal distances where the ice sheet rapidly plunges down to sea-level. Due to this abrupt transition, coarse grids will commonly represent the ablation zone with mixed grid cells containing mixtures of the ice and ice-free regions. We hypothesize that the reason that the coarser models tend to have large melt biases, despite the $ARCTCIGRIS$ having the warmest troposphere, is because summer melting is confined to these mixed ice/land/ocean grid cells. CLM deals with land heterogeneity in a complex and sophisticated manner, but CAM only see a homogenized state due to volume averaging over the sub-grid mixture. The implication is that the ice-free land patches in a grid cell may unduly influence the climate over the entire grid cell, causing a warm bias over the ice patch of a grid cell.

Figure~\ref{fig:bias} shows mean melt bias, relative to both RACMO datasets, conditionally sampled based on grid cell ice fraction in the GrIS region. Note the common ice mask approach is used to compute the errors, meaning all fields are mapped to the common ice masks and the grid cell ice fraction is defined by the common ice masks. The figure shows that coarser grids generally have two peaks in ice fraction space; a bump in positive melting errors on the 0-20$\%$ range, and another in the fully ice-covered ice cells. Also shown are the $\pm$ one standard deviation of the biases for each bin. They indicate that the biases in 0-20$\%$ bins are mostly contained in the positive bias region (a fractional bias greater than 0), whereas the fully-covered ice cells have a wider distribution, with many grid cells also containing negative melting biases. The excessive melting that occurs in 0-20$\%$ ice fraction bins supports our hypothesis that the prevalence of mixed-grid cells in representing the ablation zone in coarse grids is responsible for their large melt bias.

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=80mm]{figs/temp_xy_diffRACMO_melt.pdf}
\end{tabular}
\end{center}
\caption{Fractional melt bias over the GrIS, computed relative to the RACMO datasets using the common ice mask approach, and conditionally sampled by grid cell ice fraction provided by the common ice masks. Solid lines are the mean of the distribution with $\pm$ one standard deviation expressed by shading.}
\label{fig:bias}
\end{figure}

\subsection{Precipitation extremes}

Synoptic storms are tracked using TempestExtremes \cite{UETAL2021}. As the $ARCTIC$ grid contains $\frac{1}{4}^{\circ}$ refinement north of about $45^{\circ}$ latitude, the storm tracker is applied to this region for the $ARCTIC$ and $ne30pg3$ run to identify differences in storm characteristics due to horizontal resolution. The composite mean precipitation maps are similar between the two grids, and exhibit the iconic comma structure of synoptic cyclones (not shown).

Figure~\ref{fig:comp-pdf} is a PDF of the precipitation rates associated with the storms, by month. The PDFs are constructed by sampling all the precipitation rates within $30^{\circ}$ of the storm center, for each point on the storm track and for all storms. The PDFs are evaluated on an identical composite grid for all runs, and so storm statistics are not impacted by differences in output resolution. The $ARCTIC$ run has larger extreme precipitation rates compared to $ne30pg3$ in every month, but the increase is greatest in the summer months, which coincides with the most extreme events of the year. This is primarily due to an increase in resolution, and not the reduced physics times-step; the $ne30pg3^{*}$ run only marginally increases the extreme precipitation rates compared with $ne30pg3$ (Figure~\ref{fig:comp-pdf}).

The extreme precipitation rates in the $ARCTIC$ run are closer to the ERA5 reanalysis than in $ne30pg3$ (Figure~\ref{fig:comp-pdf}). It's difficult to know the extent that the extreme precipitation rates in ERA5 are constrained by data assimilation, or whether these precipitation rates are due to using a similar $\frac{1}{4}^{\circ}$ model as the $ARCTIC$ grid. However, it is well documented that $\frac{1}{4}^{\circ}$ models are more skillful at simulating extreme events \cite{BetAl2013JC,OETAL2016JAMES}, and so this is an additional benefit of the variable-resolution grids.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_composite_ge45N_pdf.pdf}
\end{center}
\caption{PDFs of the total precipitation rate associated with tracked storms, by month, in the $ne30pg3$, $ne30pg3^{*}$ and $ARCTIC$ runs, and compared with the ERA5 dataset.}
\label{fig:comp-pdf}
\end{figure}

\section{Conclusions}\label{sec:conclusions}

Six grids from different dynamical cores in CESM2.2 are evaluated in an AMIP-style configuration for their performance over the Arctic and in simulating the Greenland Ice Sheet (GrIS) surface mass balance (SMB). The $1-2^{\circ}$ finite-volume grids have enhanced resolution over polar regions due to the convergence of meridian lines, although a polar filter is employed to prevent atmospheric features from forming at this higher resolution. Spectral-element grids comparable to the resolution of the finite-volume grids have an isotropic grid structure, meaning the grid resolution is similar over the entire domain. Two variable-resolution grids were developed and introduced into CESM2.2 as part of this work. They use the spectral-element dycore; the $ARCTIC$ grid has $\frac{1}{4}^{\circ}$ refinement over the broader Arctic, whereas the $ARCTICGRIS$ grid is identical except it has an $\frac{1}{8}^{\circ}$ patch of refinement over Greenland.

In general, the finite-volume grids have colder summer temperatures over the Arctic compared with the spectral-element grids (incl. the variable-resolution grids). The cloud biases in all the uniform resolution grids, whether finite-volume or spectral-elements, are similar, in general being too cloudy over Arctic land masses. The variable-resolution grids largely improve the cloud biases. It should be emphasized that our analysis is specific to the Arctic summer due to its relevance to melt rates over GrIS; improved clouds in the Arctic should not be taken to mean that lower latitude regions have an improved cloud field as well.

At the regional level, a halo of negative cloud bias is found around the oceanic perimeter of Greenland in all $1-2^{\circ}$ grids, and is absent in the variable-resolution grids. This halo bias coincides with a positive cloud bias over the interior of the ice sheet. This pattern has been traced back to the inadequacy of the lower resolution grids to resolve the orographic precipitation process in Greenland. Synoptic systems moving into Greenland are not sufficiently lifted when encountering the steep ice margins due to overly smooth topography in the $1-2^{\circ}$ grids. As as result, moisture penetrates and dumps excess precipitation into the interior of the GrIS, instead of being concentrated closer to the coastal margins as indicated by observations. This results in a positive precipitation and cloud bias in the ice sheet interior, and a halo of low cloud bias about the perimeter of Greenland. The agreement of different observational data products on this bias lends confidence in attributing the cause of the precipitation and cloud biases. The variable-resolution grids compare more favorably to the observations and indicate the orographic precipitation process in Greenland is largely resolved.

The primary source and sink terms of the SMB equation are integrated over the GrIS, and evaluated in the six grids. The uniform $1-2^{\circ}$ grids all have large positive accumulation biases owing to their inability to resolve the orographic precipitation process. The uniform spectral-element grids have larger accumulation biases, suggesting that the finite-volume grids are more skillful at resolving the precipitation processes due to their finer grid spacing over Greenland, despite a polar filter. The variable resolution grids have the most accurate accumulation rates of all the grids.

The primary mass sink term of the GrIS, ice/snow melt, expresses similar biases as the accumulation rates. The uniform resolution spectral-element grids melt too much, while the finite-volume grids have reduced biases. It's more difficult to attribute these biases to grid resolution alone; the finite-volume grids have colder summers that is consistent with their lower melt bias. However, the $ARCTICGRIS$ grid has the warmest summer temperatures of all grids, yet it has reduced melting compared to the uniform resolution spectral-element grids. This suggests that grid resolution is responsible for a large fraction of the melt biases. A mechanism is proposed; that coarse grids represent the ablation zones using mixed surface-type grid cells, and the ice-free patches of the cells may largely determining the volume mean state and lead to a warm bias over the ice covered patches of the grid cell. This mechanism is supported by analysis of melt biases binned by grid cell ice fraction.

The $ARCTIC$ grid provides substantial improvements to the simulated climate of the Arctic, such as precipitation extremes and the surface mass balance over Greenland, compared with the uniform $1^{\circ}-2^{\circ}$ grids. The $ARCTICGRIS$ grid has the most realistic cloud and precipitation fields, but suffers from excessive summer temperatures. The $1^{\circ}$ finite-volume model provides a surprisingly realistic surface mass balance, likely due to the larger degrees of freedom used to represent the GrIS in latitude-longitude grids. In particular, the large number of grid cells representing the ablation zone leads to a reduction in mixed ice-covered/ice-free grid cells that seem to be so detrimental to the ablation process in the other uniform resolution grids.

As modeling systems move away from latitude-longitude grids, and towards quasi-uniform unstructured grids, it is worth taking stock as to whether this will adversely impact the skill of the simulated climate in polar regions. In our case, we have found that the $1^{\circ}$ finite-volume model has clear advantages over the $1^{\circ}$ spectral element model in simulating the surface mass balance of the GrIS. While this finding is not sufficient to slow-down the already ongoing transition towards unstructured grids in CESM, largely driven by gains in computational efficiency, it has inspired us to develop alternative configurations that recover the fidelity of polar climate in latitude-longitude grids. We have shown in this study (and a prior companion study \cite{VETAL2018TC}), that for CESM, the inclusion of Arctic refined meshes can substantially improve the simulated mass balances of the GrIS, even compared with the finite-volume $1^{\circ}$ grid. This should provide the CESM modeling community with some comfort; the ongoing transition away from latitude-longitude grids will not adversely impact CESM's reputation as a state-of-the-art tool for simulating and understanding polar processes.

We are actively working to develop a configuration of the $ARCTIC$ mesh that is fully-coupled with the CESM ocean model, and the Community Ice Sheet Model (CISM) to provide multi-century projections of the state of the GrIS and its contribution to sea-level rise. A fully-coupled pre-industrial control configuration has already been developed and vetted, and should be incorporated into CESM in the near-future. A visualization of the $ARCTCGRIS$ has been developed and made available on youtube\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}} in order to advertise some of the stunning capabilities of the refined mesh approach. A snapshot of this visualization is provided in Figure~\ref{fig:viz}, illustrating mesoscale katabatic winds descending down the southeastern slopes of GrIS. We are optimistic about the polar science opportunities provided by these new Arctic mesh configurations, and our open to collaborating with the CESM community to assist in anyway we can.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/Vis1472.jpg}
\end{center}
\caption{Snapshot of the lowest model level streamlines from the $ARCTICGRIS$ visualization, with color shading denoting the wind magnitude.}
\label{fig:viz}
\end{figure}

% \section{Materials and Methods}
% Here is text on Materials and Methods.
%
% \subsection{A descriptive heading about methods}
% More about Methods.
%
% \section{Data} (Or section title might be a descriptive heading about data)
%
% \section{Results} (Or section title might be a descriptive heading about the
% results)
%

%\begin{figure}[t]
%\begin{center}
%         \includegraphics[width=100mm]{figs/temp_tseries_BASIN_uqmap.pdf}
%\end{center}
%\caption{.}
%\label{fig:zwally}
%\end{figure}

%Text here ===>>>

%%

%  Numbered lines in equations:
%  To add line numbers to lines in equations,
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%% Enter Figures and Tables near as possible to where they are first mentioned:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables;  other caption information
%  should be placed in last line of the table, using
% \multicolumn2l{$^a$ This is a table note.}
%
%----------------
% EXAMPLE FIGURES
%
% \begin{figure}
% \includegraphics{example.png}
% \caption{caption}
% \end{figure}
%
% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
% \begin{figure}
% \noindent\includegraphics[width=\textwidth]{anothersample.png}
%\caption{caption}
%\label{pngfiguresample}
%\end{figure}
%
%
% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options. This is common when trying to add a PDF figure without pdflatex.
% \begin{figure}
% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
%\caption{caption}
%\label{pdffiguresample}
%\end{figure}
%
%
% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
%

%
% ---------------
% EXAMPLE TABLE
%
% \begin{table}
% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
% \centering
% \begin{tabular}{l c}
% \hline
%  Run  & Time (min)  \\
% \hline
%   $l1$  & 260   \\
%   $l2$  & 300   \\
%   $l3$  & 340   \\
%   $h1$  & 270   \\
%   $h2$  & 250   \\
%   $h3$  & 380   \\
%   $r1$  & 370   \\
%   $r2$  & 390   \\
% \hline
% \multicolumn{2}{l}{$^{a}$Footnote text here.}
% \end{tabular}
% \end{table}

%% SIDEWAYS FIGURE and TABLE
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{figsamp}
% \caption{caption here}
% \label{newfig}
% \end{sidewaysfigure}
%
%  \begin{sidewaystable}
%  \caption{Caption here}
% \label{tab:signif_gap_clos}
%  \begin{tabular}{ccc}
% one&two&three\\
% four&five&six
%  \end{tabular}
%  \end{sidewaystable}

%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
%\begin{linenomath*}
%\begin{equation}
%y|{f} \sim g(m, \sigma),
%\end{equation}
%\end{linenomath*}

%%% End of body of article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary, Notation or Acronym section goes here:
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
%  \begin{glossary}
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \end{glossary}

%
%%%%%%%%%%%%%%
% Acronyms
%   \begin{acronyms}
%   \acro{Acronym}
%   Definition here
%   \acro{EMOS}
%   Ensemble model output statistics
%   \acro{ECMWF}
%   Centre for Medium-Range Weather Forecasts
%   \end{acronyms}

%
%%%%%%%%%%%%%%
% Notation
%   \begin{notation}
%   \notation{$a+b$} Notation Definition here
%   \notation{$e=mc^2$}
%   Equation in German-born physicist Albert Einstein's theory of special
%  relativity that showed that the increased relativistic mass ($m$) of a
%  body comes from the energy of motion of the body—that is, its kinetic
%  energy ($E$)—divided by the speed of light squared ($c^2$).
%   \end{notation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS
%
% The acknowledgments must list:
%
% >>>>	A statement that indicates to the reader where the data
% 	supporting the conclusions can be obtained (for example, in the
% 	references, tables, supporting information, and other databases).
%
% 	All funding sources related to this work from all authors
%
% 	Any real or perceived financial conflicts of interests for any
%	author
%
% 	Other affiliations for any author that may be perceived as
% 	having a conflict of interest with respect to the results of this
% 	paper.
%
%
% It is also the appropriate place to thank colleagues and other contributors.
% AGU does not normally allow dedications.

\acknowledgments
This material is based upon work supported by the National Center for Atmospheric Research (NCAR), which is a major facility sponsored by the NSF under Cooperative Agreement 1852977. Computing and data storage resources, including the Cheyenne supercomputer (doi:10.5065/D6RX99HX), were provided by the Computational and Information Systems Laboratory (CISL) at NCAR. A. Herrington is especially grateful to W. Lipscomb (NCAR) for providing crucial feedback in his ASP postdoc application, that ultimately funded this study. A. Herrington also thanks Matt Rehme (NCAR/CISL) for his role in generating the $ARCTICGRIS$ visualization available on youtube ({\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}).

The data presented in this manuscript is available at {\url{https://github.com/adamrher/2020-arcticgrids}}.

%% ------------------------------------------------------------------------ %%
%% References and Citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% \bibliography{<name of your .bib file>} don't specify the file extension
%
% don't specify bibliographystyle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\bibliography{ enter your bibtex bibliography filename here }
\bibliography{bib}

%Reference citation instructions and examples:
%
% Please use ONLY \cite and \citeA for reference citations.
% \cite for parenthetical references
% ...as shown in recent studies (Simpson et al., 2019)
% \citeA for in-text citations
% ...Simpson et al. (2019) have shown...
%
%
%...as shown by \citeA{jskilby}.
%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
%...has been shown \cite{jskilbye}.
%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown by \cite <e.g.,>[and others]{lewin76}.
%
% apacite uses < > for prenotes and [ ] for postnotes
% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
%

\end{document}

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

% If you are using line numbers, remember to surround
% equations with \begin{linenomath*}...\end{linenomath*}

%  To add line numbers to lines in equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}